\chapter{Pack Files}
\label{chap:packfiles}

This chapter will introduce you to one file type -- The Pack File.  This file
type is used to store a large portion of the game data, and has many subtypes
for data such as animations, models, maps, textures, and audio.  You'll learn
how to navigate through the pack file to grab individual portions of the data
it holds, as well as what data each subtype contains.

\section{Headers}
\label{sec:pfheaders}

Each Pack File begins with a 12-byte header identifying the type of data
contained.  This header includes the 4-byte Character Code identifier labeling
the file as a Pack File. The format of this header can be found in
\autoref{tab:pfheader}.

\begin{datastructure}{htp}{the Pack File header}{pfheader}
	\field{char[2]}{Identifier}{Always [`P',`F']}
	\field{uint32}{Unknown}{Always \hex{01}}
	\field{uint16}{Header Size}{Always 12}
	\field{FourCC}{Type}{4 Character Codes defining the data held in this Pack File.}
\end{datastructure}

Each Pack File is split into blocks of data, called Chunks.  The first chunk
follows immediately after the Pack File header.  Each one contains a pointer
to the next one.

Chunks are identified by 4 character codes, which determines the format of the
chunk and the data stored in it.  While different Pack Files may share chunks
with the same identifier, these chunks may not be the same.  For example, the
Content Manifest and Map Metadata Pack Files both have a chunk labeled 'Main,'
but their formats are different.

Chunks begin with a 16-byte header describing where the next chunk is in the
file and the type of data stored in this chunk.  The format can be found in
\autoref{tab:pfchunkheader}.

\begin{datastructure}{htp}{the Chunk header}{pfchunkheader}
	\field{FourCC}{Identifier}{Type of Chunk}
	\field{uint32}{Next Chunk}{Number of bytes after the end of this field that%
	                           the next chunk appears}
	\field{uint16}{Unknown}{}
	\field{uint16}{Header Size}{Always 12}
	\field{Ptr\tmpl{?}}{Unknown}{Pointer to an unknown data structure}
\end{datastructure}

In \autoref{tab:pftypes} you will find all the known Pack File types, as well
as the chunks each one contains.  The following sections will go into each chunk
and its format.

\begin{table}[htp]\begin{center}
	\caption{Pack File Types sorted by FourCC}
	\label{tab:pftypes}
	
	\begin{tabular}{rrl|p{1.5in}}
		\textbf{FourCC} & \textbf{Value} & \textbf{Name} & \textbf{Included Chunks} \\
		\hline%\fourcc{`0'}{`0'}{`0'}{`0'}
		\fourcc{`A'}{`B'}{`I'}{`X'} & \hex{58494241}
			& \hyperref[sec:pfABIX]{Bank Index} %
			& \fourcc{`B'}{`I'}{`D'}{`X'} \\
		\hline
		\fourcc{`A'}{`B'}{`N'}{`K'} & \hex{4b4e4241}
			& \hyperref[sec:pfABNK]{Bank} %
			& \fourcc{`B'}{`K'}{`C'}{`K'} \\
		\hline
		\fourcc{`A'}{`M'}{`A'}{`T'} & \hex{54414d41}
			& \hyperref[sec:pfAMAT]{Material} %
			& \fourcc{`G'}{`R'}{`M'}{`T'}, \fourcc{`D'}{`X'}{`9'}{`S'} \\
		\hline
		\fourcc{`A'}{`M'}{`S'}{`P'} & \hex{50534d41}
			& \hyperref[sec:pfAMSP]{Script} %
			& \fourcc{`A'}{`M'}{`S'}{`P'} \\
		\hline
		\fourcc{`a'}{`n'}{`i'}{`c'} & \hex{63696E61}
			& \hyperref[sec:pfanic]{Animation Sequence} %
			& \fourcc{`s'}{`e'}{`q'}{`n'} \\
		\hline
		\fourcc{`A'}{`S'}{`N'}{`D'} & \hex{444e5341}
			& \hyperref[sec:pfASND]{Audio} %
			& \fourcc{`A'}{`S'}{`N'}{`D'} \\
		\hline
		\fourcc{`C'}{`I'}{`N'}{`P'} & \hex{504e4943}
			& \hyperref[sec:pfCINP]{Scene} %
			& \fourcc{`C'}{`S'}{`C'}{`N'} \\
		\hline
		\fourcc{`c'}{`m'}{`a'}{`C'} & \hex{43616d63}
			& \hyperref[sec:pfcmaC]{Collide Model Manifest} %
			& \fourcc{`m'}{`a'}{`i'}{`n'} \\
		\hline
		\fourcc{`c'}{`m'}{`p'}{`c'} & \hex{63706d63}
			& \hyperref[sec:pfcmpc]{Composite} %
			& \fourcc{`c'}{`o'}{`m'}{`p'} \\
		\hline
		\fourcc{`c'}{`n'}{`t'}{`c'} & \hex{63746e63}
			& \hyperref[sec:pfcntc]{Content Manifest} %
			& \fourcc{`M'}{`a'}{`i'}{`n'} \\
		\hline
		\fourcc{`e'}{`m'}{`o'}{`c'} & \hex{636f6d65}
			& \hyperref[sec:pfemoc]{Emote Animation} %
			& \fourcc{`a'}{`n'}{`i'}{`m'} \\
		\hline
		\fourcc{`e'}{`u'}{`l'}{`a'} & \hex{616c7565}
			& \hyperref[sec:pfeula]{EULA} %
			& \fourcc{`e'}{`u'}{`l'}{`a'} \\
		\hline
		\fourcc{`h'}{`v'}{`k'}{`C'} & \hex{436b7668}
			& \hyperref[sec:pfhvkC]{Havok} %
			& \fourcc{`h'}{`a'}{`v'}{`k'} \\
		\hline
		\fourcc{`m'}{`a'}{`p'}{`c'} & \hex{6370616d}
			& \hyperref[sec:pfmapc]{Map} %
			& \fourcc{`a'}{`u'}{`d'}{`i'},    \fourcc{`m'}{`s'}{`n'}{\hex{00}},
			\fourcc{`p'}{`a'}{`r'}{`m'},      \fourcc{`s'}{`h'}{`o'}{`r'},
			\fourcc{`s'}{`u'}{`r'}{`f'},      \fourcc{`t'}{`r'}{`n'}{`i'},
			\fourcc{`a'}{`r'}{`e'}{`a'},      \fourcc{`h'}{`a'}{`v'}{`k'},
			\fourcc{`c'}{`u'}{`b'}{`e'},      \fourcc{`d'}{`c'}{`a'}{`l'},
			\fourcc{`e'}{`n'}{`v'}{\hex{00}}, \fourcc{`l'}{`g'}{`h'}{`t'},
			\fourcc{`p'}{`r'}{`p'}{`2'},      \fourcc{`r'}{`i'}{`v'}{`e'},
			\fourcc{`s'}{`h'}{`e'}{`x'},      \fourcc{`t'}{`r'}{`n'}{\hex{00}},
			\fourcc{`z'}{`o'}{`n'}{`2'} \\
		\hline
		\fourcc{`m'}{`M'}{`e'}{`t'} & \hex{74654d6d}
			& \hyperref[sec:pfmMet]{Map Metadata} %
			& \fourcc{`M'}{`a'}{`i'}{`n'} \\
		\hline
		\fourcc{`M'}{`O'}{`D'}{`L'} & \hex{4c444f4d}
			& \hyperref[sec:pfMODL]{Model} %
			& \fourcc{`A'}{`N'}{`I'}{`M'}, \fourcc{`M'}{`O'}{`D'}{`L'},
			\fourcc{`G'}{`E'}{`O'}{`M'}, \fourcc{`P'}{`R'}{`P'}{`S'},
			\fourcc{`R'}{`O'}{`O'}{`T'}, \fourcc{`S'}{`K'}{`E'}{`L'} \\
		\hline
		\fourcc{`m'}{`p'}{`s'}{`d'} & \hex{6473706d}
			& \hyperref[sec:pfmpsd]{Map Shadow} %
			& \fourcc{`s'}{`h'}{`a'}{`d'} \\
		\hline
		\fourcc{`P'}{`I'}{`M'}{`G'} & \hex{474d4950}
			& \hyperref[sec:pfPIMG]{Paged Image} %
			& \fourcc{`P'}{`G'}{`T'}{`B'} \\
		\hline
		\fourcc{`p'}{`r'}{`l'}{`t'} & \hex{746c7270}
			& \hyperref[sec:pfprlt]{Portal Manifest} %
			& \fourcc{`m'}{`f'}{`s'}{`t'} \\
		\hline
		\fourcc{`t'}{`x'}{`t'}{`m'} & \hex{6d747874}
			& \hyperref[sec:pftxtm]{Text Manifest} %
			& \fourcc{`t'}{`x'}{`t'}{`m'} \\
		\hline
		\fourcc{`t'}{`x'}{`t'}{`V'} & \hex{56747874}
			& \hyperref[sec:pftxtV]{Text Variant} %
			& \fourcc{`v'}{`a'}{`r'}{`i'} \\
		\hline
		\fourcc{`t'}{`x'}{`t'}{`v'} & \hex{76747874}
			& \hyperref[sec:pftxtv]{Text Voice} %
			& \fourcc{`t'}{`x'}{`t'}{`v'} \\
		\hline
	\end{tabular}
\end{center}\end{table}

\section{Common Structures}
\label{sec:pfcommonstructures}
Before we go into the different Pack File types and their formats, we should go
over some common structures found in the pack files.  There are three basic
structures and three others that build off of them.

We'll start with the Pointer.  In the documentation, we'll shorten it to Ptr for
brevity.  This structure consists of one 32-bit offset that references data
elsewhere in the file.  The address of the data referenced is the address of the
Ptr plus the offset.

The next basic structure is the Array -- shortened to Arr.  It is simply one
data type T repeated N times.  It will be written as Arr\tmpl{T,N}, T being the
data type and N being the number of data types.  It takes up $size(T)*N$ bytes.

The last basic structure is the File Reference -- shortened to FileRef.  It
occupies 6 bytes, but only 4 of those bytes are useful information.  According
to Johan Sköld, this structure represents a File ID, and it is only valid if the
first two 16-bit values are both above \hex{100}.  In his project, he calculates
the File ID as $first - \hex{FF} + (second - \hex{100}) * \hex{FF00}$.
\emph{More research must be done into whether this equation is correct.}

Moving on to composite structures, the first we'll discuss is the String and
Wide String -- shortened to Str and WStr, respectively.  These are simply Ptrs
to C Strings using either traditional or wide characters, but are common enough
that they warrant their own listing here.

The next composite structure is the Vector -- shortened to Vtr.  This is similar
to an array, but the size is determined at run time and not at compile time, and
it always takes up 8 bytes regardless of the number of elements it represents.
The first 32-bit value in this structure, N, is the number of elements in the
Vector. The second is a Pointer of type Ptr\tmpl{Arr\tmpl{T,N}}, where T is the
data type represented by the Vector.

The last composite structure we'll discuss is the Reference List -- shortened
to RefList.  This is a Vtr of pointers to data types.  It is equivalent to
Vtr\tmpl{Ptr\tmpl{T}}, where T is the data type we're pointing to.

\input{PackFileTypes/anic}
\input{PackFileTypes/ASND}
\input{PackFileTypes/ABNK}
\input{PackFileTypes/ABIX}
\input{PackFileTypes/cmaC}
\input{PackFileTypes/cmpc}
\input{PackFileTypes/cntc}
\input{PackFileTypes/emoc}
\input{PackFileTypes/eula}
\input{PackFileTypes/hvkC}
\input{PackFileTypes/mapc}
\input{PackFileTypes/mMet}
\input{PackFileTypes/mpsd}
\input{PackFileTypes/AMAT}
\input{PackFileTypes/MODL}
\input{PackFileTypes/PIMG}
\input{PackFileTypes/prlt}
\input{PackFileTypes/CINP}
\input{PackFileTypes/AMSP}
\input{PackFileTypes/txtm}
\input{PackFileTypes/txtV_txtv}

